Toward single particle reconstruction without particle picking: Breaking the detection limit

### 2018-08-02 TEX/draft_biology.tex
removed false statement on EM

As a proof of concept, we study a simplified model for 3-D reconstruction directly from the micrograph on a simulated data. 
In order to recover the 3-D volume, we use autocorrelation analysis. In a nutshell, we relate the autocorrelations of the micrographs to the autocorrelations of the volume.
For any noise level, these autocorrelations can be estimated to any desired accuracy, provided the we get to observe sufficiently many particle projections and the latter are separated in the micrograph. Importantly, there is no need to detect the projections. The autocorrelations of the micrographs are straightforward to compute and require only one pass over the data. These directly yield estimates for the autocorrelations of the volume. To estimate the volume itself from its estimated autocorrelations, we solve a nonlinear inverse problem via least-squares. In this controlled experiment, we make several simplified assumptions; see the Discussion Section for details. 
As a side note, we mention that expectation-maximization (EM)---a popular framework in SPR---is intractable for this problem; see~\cite{bendory2018estimation} for the details. \TODO{Do we want to stress that we get only low-resolution of the volume?}

@@

We represent our autocorrelations using Prolate Spheroidal Wave Functions (PSWFs) $\{\psi_{k,q}\}$ where $k \geq 0, q \geq 1$ are integers~\cite{slepian1964pswfIV}. As we demonstrate below, this makes it easier to account for the fact that all in-plane rotations of the micrographs are equally likely observations. This is only a concern for the second and third order autocorrelations. Below, we start with $p=2$.
The PSWFs are eigenfunctions of the truncated Fourier transform and are given in polar coordinates by
\begin{align}
	\psi_{k,q}(r,\varphi) & = \left\{\begin{array}{ll} \frac{1}{\sqrt{8\pi}}\alpha_{k,q}R_{k,q}(r)e^{\iota k\varphi}, & r\leq 1,\\ 0, & r>1,\end{array}\right. \label{eq:prolatesdef}
\end{align}
where the ${R_{k,q}}$ are a family of real, one-dimensional functions and the ${\alpha_{k,q}}$ are \TODO{real?} scaling factors which will be defined in the next section.
The PSWFs are orthonormal \TODO{orthogonal?} on the unit disk.

>>>>

As a proof of concept, we study a simplified model for 3-D reconstruction directly from the micrograph on a simulated data. 
In order to recover the 3-D volume, we use autocorrelation analysis. In a nutshell, we relate the autocorrelations of the micrographs to the autocorrelations of the volume.
For any noise level, these autocorrelations can be estimated to any desired accuracy, provided the we get to observe sufficiently many particle projections and the latter are separated in the micrograph. Importantly, there is no need to detect the projections. The autocorrelations of the micrographs are straightforward to compute and require only one pass over the data. These directly yield estimates for the autocorrelations of the volume. To estimate the volume itself from its estimated autocorrelations, we solve a nonlinear inverse problem via least-squares. In this controlled experiment, we make several simplified assumptions; see the Discussion Section for details. 
%As a side note, we mention that expectation-maximization (EM)---a popular framework in SPR---is intractable for this problem; see~\cite{bendory2018estimation} for the details. 
\TODO{Do we want to stress that we get only low-resolution of the volume?}

@@

We represent our autocorrelations using Prolate Spheroidal Wave Functions (PSWFs) $\{\psi_{k,q}\}$ where $k \geq 0, q \geq 1$ are integers~\cite{slepian1964pswfIV}. As we demonstrate below, this makes it easier to account for the fact that all in-plane rotations of the micrographs are equally likely observations. This is only a concern for the second and third order autocorrelations. Below, we start with $p=2$.
The PSWFs are eigenfunctions of the truncated Fourier transform and are given in polar coordinates by
\begin{align}
	\psi_{k,q}(r,\varphi) & = \left\{\begin{array}{ll} \frac{1}{\sqrt{8\pi}}\alpha_{k,q}R_{k,q}(r)e^{\iota k\varphi}, & r\leq 1,\\ 0, & r>1,\end{array}\right. \label{eq:prolatesdef}
\end{align}
\TODO{Ref eq. 8 in Boris' paper}
where the ${R_{k,q}}$ are a family of real, one-dimensional functions and the ${\alpha_{k,q}}$ are  scaling factors which will be defined in the next section.
The PSWFs are orthogonal on the unit disk.
